\documentclass{article}

\usepackage[english]{babel}
\usepackage[letterpaper,top=2.5cm,bottom=2.5cm,left=2.5cm,right=2.5cm,marginparwidth=1.75cm]{geometry}
\usepackage{amsmath, fancyhdr, graphicx, tikz, pgfplots}
\begin{document}
For each lab experiment you have to write an engineering report comprising/analyzing/concluding engineering design parameters from the data collected. The report must have the following format and include the following parts.
\section{Cover Page}
    Containing the course number, lab, department, school, title, your name, and the date. Ensure a blank page follows the cover page.
\section{Purpose of the Experiment}
    One paragraph briefly describing procedures and parameters being investigated.
\section{Procedures and Equipment}
    This will be brief you just need to show that you need to include it so that people are able to reproduce the experiment.
\section{Data and Results}
    This section contains tables of calculated numbers and actual numbers from the data based on the formulas collected along with unit conversions. Show graphs of any of these to see correlations.
\section{Sample Calculations}
    One-to-two pages of sample calculations, one of each to show how you have manipulated the data and the results.
\section{Discussion}
    Two parts. The first part is theory which is any theoretical subject (briefly) relating to the experiment. The next part is the experiment where you discuss in detail whatever trends you see or conclude from the tables and graphs where you have presented the data and results.
\section{Conclusions}
    You have to present, with numbers, all the important items you have presented in the experiment discussion. Nothing you've read somewhere. No general comments that cannot be seen from the data and results discussion.
\section{Questions}
    Answer all of the questions posed on the laboratory procedure sheet. Some of these answers can also be used in the discussion.
\section{Raw Data}
    In many instances, class data will be included with individual data. Include the data collected in the lab during the course of the experiment.
\section{References}
    List any references you have used. Don't forget Professor Yapijakis's lecture notes.
\end{document}