\part*{Week 4}
\addcontentsline{toc}{part}{Week 4}
\section*{Chapter 5}
\underline{9-14}
\addcontentsline{toc}{section}{Chapter 5: 9-14}
\subsection*{Problem 9}
How is monitoring conducted for lead and copper in drinking water? What are the action levels, and what are the treatment techniques if these action levels are exceeded?
\\\rule{5cm}{1pt}
\\\\Monitoring is conducted for lead and copper in first-flush tap samples. The action level for lead is 0.015 mg/L and the action level for copper is 1.3 mg/L. The treatment technique is to perform corrosion studies. Additionally, a public education program on the many health risks and actions that can be taken to reduce water pollution is required.
\subsection*{Problem 10}
What is the optimum concentration of fluoride in drinking water in a location where the average maximum daily air temperature is 60°F? What is the health benefit of drinking water containing the optimum concentration of fluoride?
\\\rule{5cm}{1pt}
\\\\The optimum concentration of fluoride in drinking water in a location where the maximum daily air temperature is 60°F is 1 mg/L. The optimum concentration of fluoride in drinking water prevents teeth from decay. 
\subsection*{Problem 11}
What is the health risk of nitrate in drinking water?
\\\rule{5cm}{1pt}
\\\\The health risk fo nitrate in drinking water is Methemoglobinemia in infants 3 to 6 months of age.
\subsection*{Problem 12}
What are the most frequently detected VOCs in contaminated groundwater? What pesticide SOCs have been detected in groundwater?
\\\rule{5cm}{1pt}
\\\\The most frequently detected VOCs in contaminated groundwater are \textit{trichloroethylene}, \textit{tetrachloroethylene}, \textit{carbon tetrachloride}, \textit{1,1, 1-trichloroethane}, \textit{1, 2-dichloroethane}, and \textit{vinyl chloride}. THe pesticide SOCs that have been detected in groundwater are \textit{alachlor}, \textit{aldicarb}, \textit{atrazine}, \textit{carbofuran}, \textit{ethylene dibromide}, and \textit{dibromochloropropane}.
\newpage
\subsection*{Problem 13}
Write the chemical formulas for the five trihalomethane compounds that have been found in drinking water. What is the source of the THMs?
\\\rule{5cm}{1pt}
\begin{enumerate}
    \item \(\text{CHCl}_3\)
    \item \(\text{CHBrCl}_2\)
    \item \(\text{CHBr}_2\text{Cl}\)
    \item \(\text{CHBr}_3\)
    \item \(\text{CHCl}_2\text{I}\)
\end{enumerate}
The source of THMs is the treatment of surface waters, particularly with chlorination. 
\subsection*{Problem 14}
Why are iron and manganese included in secondary standards for aesthetics?
\\\rule{5cm}{1pt}
\\\\Iron and manganese are included in secondary standards for aesthetics due to the brownish-colored staining of laundry and porcelain.

